---
title: "holeyness_analysis_maybe_drop"
author: "Patrick Alexander Walkden"
date: "11/01/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load in packages and data, echo = FALSE}
rm(list = ls())

## loading in tidyverse and the functions I have created to analyise and visualise the differences between TPD hypervolumes
require(tidyverse)#
require(ggpubr) # to aggregate plots
require(lme4)
require(ggResidpanel)
require(multcomp)
source("../Functions/TPD_3D_plots.R")

## load in all the data

 PREDICTS_tpds <- readRDS("../Outputs/PREDICTS_sites_tpds.rds") # morphometric TPDs of observed sites
 
 PREDICTS_randomisations <- readRDS("../Outputs/randomisations_TPD_morpho.rds") ## morphometric TPD of randomisised sites 
 PREDICTS_randomisations_site_ranges <- readRDS("../Outputs/randomisations_TPD_morpho_site_ranges.rds")
 
 
 PREDICTS_full <- readRDS("../Outputs/refined_predicts.rds")
 
 PREDICTS <- PREDICTS_full %>%  ## PREDICTS data 
   dplyr::distinct(SSBS, Predominant_habitat, Use_intensity, Biome, UN_subregion, Realm, SS) %>% ## pull out land_use type, Subregion, realm etc
   dplyr::mutate(Predominant_habitat = ifelse(grepl(Predominant_habitat, pattern = "secondary", ignore.case = TRUE), "Secondary vegetation",paste(Predominant_habitat)), 
                 Predominant_habitat = ifelse(grepl(Predominant_habitat, pattern = "primary", ignore.case = TRUE), "Primary vegetation",paste(Predominant_habitat))) %>% data.frame() ## merge all secondary sites together
 
 species_TPD <- readRDS("../Outputs/species_tpds_morpho.rds") ## individual species TPDs 
 
 Forage <- readRDS(file = "../../PREDICTS_Taxonomy/PREDICTS_imputed_BL_traits_forage.rds")

 
 species_pools <- readRDS("../Outputs/predicts_sites_species_pools.rds")


```

# site distribution

First things first, lets see how the sites that have a TPD calculated for them are distributed across land-uses and biogeographic realms.


```{r sites in lu and realm}
 
### get the land uses and use intensity of all the sites TPDs could be caluclated for 

 TPD_LU <- data.frame(SSBS = names(PREDICTS_tpds)) %>% dplyr::left_join(PREDICTS) %>% dplyr::filter(Predominant_habitat != "Cannot decice",
                                                                                                    Use_intensity != "Cannot decide")
### check a table
 table(TPD_LU$Predominant_habitat, TPD_LU$Realm)
```
I have decided to pool together sites with biogeographic realms as

Here we can see that there is quite an uneven spread of sites across the LU types and Realms so to check whether there are enough sites I am going to iteratively calculate functional diversity metrics for each LU-realm combination and if FD reaches an asymptote it will indicate that there are sufficent sites so that trait spaces can be compared.

```{r LU realm combos, echo = FALSE, cache=TRUE, message=FALSE, results = 'hide'}

### pull land_uses

land_uses <- PREDICTS %>% dplyr::distinct(Predominant_habitat, .keep_all = FALSE) %>% dplyr::filter(Predominant_habitat != "Cannot decide") %>% 
   pull() %>% as.character()

## pull realms 

realms <- as.character(unique(TPD_LU$Realm))
   
### get combinations of land uses and realms

combinations <- expand.grid(land_uses,realms) %>% set_colnames(c("land_use","realm"))

## check list for each combination I am adding each site in turn and calculating the resultant Functional Richness of the overall hypervolume - if the functional richness asymptotes before the sites run out then there are enough samples for the LUs to be compared  

if(any(grepl(list.files("../Outputs"), pattern = "check_list"))){
 Check_list <- readRDS(grep(list.files("../Outputs", full.names = TRUE), pattern = "check_list", value = TRUE)) 
} else {

  Check_list <- list()
  for(i in 1:nrow(combinations)){
  
    Land_use <- as.character(combinations[i,"land_use"])
    Realm <- as.character(combinations[i,"realm"])
    
    # This function will compute these metric and also produce a ggplot representation
    
    list <-   TPD_site_check(data = PREDICTS_tpds, LU = Land_use , realm = Realm)
    
    Check_list[[i]] <- list
    names(Check_list)[i] <- paste(Land_use, Realm, sep = "/")
  }
  write_rds("../Outputs/realm_lu_check_list.rds", x = Check_list)
}
```
Now let's have a look! I have checked all the plot but I am going to exhibit just a couple here 

```{r realm checking}
sites <- c(1,11,31)

for(i in sites){
  
  if(is_empty(Check_list[[i]])){
    next()
  }
    figure <- Check_list[[i]]$ggplots
    figure <- annotate_figure(figure,
                    top = text_grob(paste(names(Check_list)[i]), color = "red", face = "bold", size = 14))

plot(figure)
}


```

So upon examining the plots there are a good number of LU/realm combinations that have sufficient sites to reasonably represent the occupancy of functional trait space.

Above are some examples of the variety of different patterns produced. Plantation forest in the Palearctic clearly is sufficeintly sampled. Pasture in the Afrotropics. although containing only a few sites does exhibit an asymptote and has been included for now but may need to be dropped later, and finally cropland in the Nearctic which is showing Functional richness is still increasing so therefore has not been sufficeintly sampled and therefore dropped. 

So....

In the Afrotropics we can compare Cropland, Pasture, Plantation forest, Primary forest, Primary non-forest and Secondary vegetation.

In Australasia we can compare Pasture, Primary forest, Primary non-forest and Secondary vegetation.

In Indo-Malay we can compare Cropland, Plantation forest, Primary forest, and Secondary vegetation.

In the Nearctic we can compare Cropland, Pasture, Primary forest, Primary non-forest and Secondary vegetation.

In the Neotropics we can compare Cropland, Pasture, Plantation forest, Primary forest, and Secondary vegetation.

In the Palearctic we can compare Cropland, Pasture, Plantation forest, Primary forest, Primary non-forest and Secondary vegetation.


Primary forest is the only land-use classification that is well represented ubiquitously across the realms, but there is also good coverage from secondary vegetation, plantation forest, pasture and, maybe, cropland. 

Urban and primary non-forest classifications are represented well in only three of the realms... maybe a case for combining primary sites together??

However, for now lets have a look at these trait spaces and use Primary forest sites in the Afrotropics as an example.


```{r site occupancy visualisation, webgl = TRUE}

pri_afro <- TPD_LU %>% dplyr::filter(Predominant_habitat == "Primary vegetation", Realm == "Afrotropic") %>% dplyr::distinct(SSBS) %>% pull()

TPD_3d_plot(data = PREDICTS_tpds, sites = pri_afro, T1lab = "Locomotion", T2lab = "Foraging", T3lab = "Body", method = "prob", title = "TPD_test_plot")

```

In this plot you can see a 3D plot representing the occupancy of trait space in the three dimensions characterised by our axes of the forgaing, locomotory and body niches. 

Scores on the body dimension track the size of the bird. 

High scores on the locomotory axis represent long tarsus' with slight increases in secondary wing lengths, while lower scores indicate longer tails and minor increases in wing length. 

The foraging niche is based on the dimensions of the beak and high scores representing wide, deep beaks and as scores decrease beaks tend towards those that are longer and thinner. 

It's all well and good being able to represent these trait spaces but what do they represent and how do they differ across land use gradients within biogeographic realms.

```{r FD_TPD metrics, fig.width= 10}

## create a list with the land use classifications for each realm

realm_land_uses <- list()
realm_land_uses[["Afrotropic"]] <- c("Primary vegetation", "Secondary vegetation","Plantation forest","Pasture","Cropland")
realm_land_uses[["Australasia"]] <- c("Primary vegetation", "Secondary vegetation", "Pasture")
realm_land_uses[["Indo-Malay"]] <- c("Primary vegetation","Secondary vegetation","Plantation forest","Cropland")
realm_land_uses[["Nearctic"]] <- c("Primary vegetation", "Pasture","Cropland", "Urban")
realm_land_uses[["Neotropic"]] <- c("Primary vegetation","Secondary vegetation","Plantation forest","Pasture","Cropland")
realm_land_uses[["Palearctic"]] <- c("Primary vegetation","Secondary vegetation", "Plantation forest","Pasture","Cropland","Urban")

```


### What areas of trait space are 


## Hypervolume holes


```{r hypervolume holes}

## create empty data frame to add metrics of interior and exterior holes 

holes_frame <- c()

for(r in realms){
  
  ### get the land uses 
    
  land_uses <- realm_land_uses[[r]]
  
  ## get proportion and volume of interior and exterior holes 
  hole_mat <- matrix(rep(NA,length(land_uses)*4), ncol = 4) %>% data.frame()
  rownames(hole_mat) <- land_uses
  colnames(hole_mat) <- c("total_hole_volume_int","proportion_to_observed_int","total_hole_volume_ext","proportion_to_observed_ext" )
  
  #iterate through land uses 
  
  for(LU in land_uses){
  
  lu_sites <- TPD_LU %>% dplyr::filter(Predominant_habitat == LU, Realm == r) %>% dplyr::distinct(SSBS) %>% pull()
  
  ## calculate hole metrics with a threshold of 0.95 quantile from the randomisations to be considered a hole 
  
  holes <- TPD_holes(data = PREDICTS_tpds, randata = PREDICTS_randomisations, sites = lu_sites, threshold = 1)
  
  ### add the internal and external metrics to the data frame
  
  hole_mat[LU,] <- c(unlist(holes$internal[c("total_hole_volume","proportion_holes_volume")]),
                     unlist(holes$external[c("total_hole_volume","proportion_holes_volume")]))
  }
  
  ### add columns for realm and LU
  
  hole_mat$Realm <- r
  hole_mat$Land_use <- land_uses
  hole_mat$Land_use <- factor(hole_mat$Land_use, levels = land_uses)
  rownames(hole_mat) <- 1:nrow(hole_mat)
  
  
  holes_frame <- rbind(holes_frame,hole_mat)  
  
}


holes_frame$Land_use <- factor(holes_frame$Land_use, levels = c("Primary vegetation", "Secondary vegetation", "Plantation forest", "Pasture", "Cropland", "Urban"))
## for each create a plot

for(r in realms ){
 hole_plot_data <- holes_frame %>% dplyr::filter(Realm == r)
  
  p_rich <- ggplot(data = hole_plot_data, aes(x = Land_use, y = proportion_to_observed_int)) +
    geom_bar(stat = "identity") +
    ggtitle(paste(r))
    
plot(p_rich)
}
 



```

## Where are these holes? What species occupy the empty space?

By looking at what species occupy the functional trait space left vacant in observed communities compared to randomised null communites we can begin to see how the change in the structure of functional trait space --- Could we have basically just done this with species identities?? The whole point is to get beyond this point just considering traits, although this could then be like these could be candidate species for re-introductions as they occupy niche space that is currently not being filled. 

Additionally, we can have a look better at which areas of traits space are favoured or lost across a land use gradient by looking at the differences between the hypervolumes of each last use type and identifying areas that drop in and out of occupancy.

```{r species in the holes, fig.width=10}

## First load in Foraging guild and species pools data 


## if holey species list identifying which species occupy holes left compared to random across realms and LU types load it otherwise...




pool_sp <- function(df,sites, pool){
  
  pool_species <- c()
    for(sit in sites){
      pool_species <- unique(c(pool_species,species_pools[[sit]][["0.9"]]))
    }
  
    for(i in 1:nrow(df)){
        sp <- unlist(strsplit(df[i,"occupying_species"],split = "/"))
        df[i,"occupying_species"] <- paste(sp[which(sp %in% pool_species)], collapse = "/")
    }
  
  return(df)
  
}

split_sp_func <- function(df){
  species <- c()
  for(i in 1:nrow(df)){
  species <- unique(c(species,unlist(strsplit(df[i,"occupying_species"], split = "/"))))
  }
  return(species)
}


################################################
#################################################

if(any(grepl(x = list.files("../Outputs"), pattern = "holey_species"))){
  holey_species <- readRDS(grep(list.files("../Outputs", full.names = TRUE), value = TRUE, pattern = "holey_species"))
} else {
  
# create empty list   
  
holey_species <- list()

#iterate through realms

for(r in realms){
  
  ## get the land uses 
  
  land_uses <- realm_land_uses[[r]]
  
  # iterate through land uses 
  
  for(LU in land_uses){
  
  ## get sites 
    
  lu_sites <- TPD_LU %>% dplyr::filter(Predominant_habitat == LU, Realm == r) %>% dplyr::distinct(SSBS) %>% pull()
  
  ## identify where holes are compared to random
  
  holes <- TPD_holes(data = PREDICTS_tpds, randata = PREDICTS_randomisations, sites = lu_sites[1], threshold = 0.95)
  
  ## determine which species fit in these holes based on their functional traits 
  
  internal_species <- species_fit(holes[["internal"]][["internal_hole_cells"]])
  
  ## This produces a data frame
  
  
  
  ##### now this is all the species that are in the entirety of the assembly species pool so now I'm going to filter just for those species that are in the species pools of all the sites combined.
 
  internal_species <- pool_sp(internal_species)
  
  
   holey_species[paste(r,LU,sep = "_")] <- internal_species
  }
}

write_rds(holey_species, file = "Outputs/holey_species.rds")

}









hole_plot_sp <- c()

for( r in realms){

  land_uses <- realm_land_uses[[r]]
  
for(LU in land_uses){

  data <- holey_species[[paste(r,LU,sep = "_")]]
  
r_lu_hole_sp <- split_sp_func(data)



hole_sp_for <- data.frame(Birdlife_Name = r_lu_hole_sp, Land_use = LU, Realm = r) %>% dplyr::left_join(Forage[,c("Birdlife_Name", "Trophic_Level", "Trophic_Niche", "Foraging_Niche")], by = "Birdlife_Name")

hole_sp_for$Trophic_Niche <- factor(hole_sp_for$Trophic_Niche)


hole_plot_sp <- rbind(hole_plot_sp,hole_sp_for)

}
}

hole_plot_sp$Land_use <- factor(hole_plot_sp$Land_use,
                                levels = c("Primary forest", "Primary non-forest", "Secondary vegetation", "Plantation forest",
                                           "Pasture", "Cropland", "Urban"))
hole_plot_sp$Trophic_Niche <- factor(hole_plot_sp$Trophic_Niche, levels = c("In", "Fr","Gr", "Ne", "Hb.T", "Hb.A", "Vt", "Aq.p","Om" ))

hole_plot_sp <- hole_plot_sp %>% dplyr::mutate(Foraging_Niche = ifelse(Foraging_Niche == "For.Gen", "For.Gen","Spec"))




for(r in unique(realms)){

  ### Or could weight it by the size of the hole?? But I've gone for size of the observed communitity and this point but hole size may be better 
  
  weights <- holes_frame%>% dplyr::filter(Realm == r)
  
  data <- hole_plot_sp %>% dplyr::filter(Realm == r) %>% group_by(Trophic_Niche, Foraging_Niche, Land_use) %>% dplyr::summarise(count = n(), .groups = "drop") %>% dplyr::left_join(weights[,c("Land_use", "total_hole_volume_int")], by = "Land_use") %>% dplyr::mutate(relative_sp_occupancy = count/total_hole_volume_int)
    

hole_sp_plot <- ggplot(data = data, aes( x = Land_use, y = relative_sp_occupancy, fill = Foraging_Niche)) +
  geom_bar(stat = "identity") +
  facet_grid(. ~Trophic_Niche) +
  ggtitle(paste(r,sep = "_")) +
  theme(axis.text.x = element_text(angle = 90))

plot(hole_sp_plot)
}
###################################################
###################################################
 ## Let's now look at individual site holeyness
```


```{r}
require(future)
require(future.apply)
require(car)
require(lmerTest)
require(merTools)


if(any(grepl(list.files("../Outputs"), pattern = "site_holes.rds"))){
  site_hole <- readRDS(grep(list.files("../Outputs",full.names = TRUE),pattern = "site_holes.rds", value = TRUE))
} else{

registerDoParallel(cores = 8)

all_sites <- TPD_LU$SSBS


site_holes <- foreach(site = all_sites,
                      .combine = "rbind",
                      .packages = c("tidyverse","fastcluster","geometry","magrittr","gstat" ),
                      .inorder = FALSE) %dopar% {

                      data <- TPD_LU %>% dplyr::filter(SSBS == site)  
                      
                      site_hole <- TPD_holes(data = PREDICTS_tpds, randata = PREDICTS_randomisations, sites = site, threshold = 1)

                      if(is_empty(site_hole)){
                        data$internal_hole_vol <- NA
                        data$internal_hole_prop <- NA
                        
                        data$external_hole_vol <- NA
                        data$external_hole_prop <- NA
                          
                      }else{
                      
                      data$internal_hole_vol <- site_hole$internal$total_hole_volume
                      data$internal_hole_prop <- site_hole$internal$proportion_holes_volume
                      
                      data$external_hole_vol <- site_hole$external$total_hole_volume
                      data$external_hole_prop <- site_hole$external$proportion_holes_volume
                      
                      }
                      return(data)

}

registerDoSEQ()
closeAllConnections()


site_hole <- readRDS("Outputs/site_holes.rds")
}


site_hole$Predominant_habitat <- relevel(factor(site_hole$Predominant_habitat), ref = "Primary vegetation")
site_hole$Realm <- relevel(factor(site_hole$Realm), ref = "Neotropic")





site_hole <- site_hole %>% dplyr::left_join(PREDICTS[,c("SSBS", "SS")], by = "SSBS") %>% distinct(SSBS, .keep_all = TRUE)

FD_mets <- TPD_FD_metrics(data = PREDICTS_randomisations_site_ranges, sites = site_hole$SSBS)

site_hole$F_Rich <- FD_mets$FRich
hist(site_hole$F_Rich)
hist(log(site_hole$F_Rich))

hist(site_hole$internal_hole_prop)
hist(sqrt(site_hole$internal_hole_prop))
hist(car::logit(site_hole$internal_hole_prop))

     
     

site_mod <- lmer(sqrt(internal_hole_prop) ~ Predominant_habitat + Realm + Predominant_habitat:Realm + (1|SS), data = site_hole)



summary(site_mod)
Anova(site_mod, type = "II")

ggResidpanel::resid_panel(site_mod)




newdata <- data.frame(Realm = c(rep(names(realm_land_uses)[1], length(realm_land_uses[[1]])),
                                 rep(names(realm_land_uses)[2], length(realm_land_uses[[2]])),
                                 rep(names(realm_land_uses)[3], length(realm_land_uses[[3]])),
                                 rep(names(realm_land_uses)[4], length(realm_land_uses[[4]])),
                                 rep(names(realm_land_uses)[5], length(realm_land_uses[[5]])),
                                 rep(names(realm_land_uses)[6], length(realm_land_uses[[6]]))),
                       Predominant_habitat = c(realm_land_uses[[1]],
                                    realm_land_uses[[2]],
                                    realm_land_uses[[3]],
                                    realm_land_uses[[4]],
                                    realm_land_uses[[5]],
                                    realm_land_uses[[6]]))



#for(i in 1:nrow(lala)){
#  ph <- lala[i,"Predominant_habitat"]
#  r <- lala[i,"Realm"]
#lala[i,"logF_Rich"] <- log(mean(site_hole[which(site_hole$Predominant_habitat == ph & site_hole$Realm == r),"F_Rich"]))
#}



pred_fun <- function(x){
  as.numeric(predict(x, newdata = newdata, re.form = NA))
}

newdata$predicted <- pred_fun(site_mod)^2
newdata$upper <- NA
newdata$lower <- NA


booted_mod <- bootMer(site_mod, FUN = function(x) pred_fun(x), nsim = 1000)

estimates <- booted_mod$t0^2

bootstraps <- booted_mod$t^2

for(i in 1:ncol(bootstraps)){
  newdata[i,"upper"] <- as.numeric(quantile(bootstraps[,i], 0.95))
  newdata[i,"lower"] <- as.numeric(quantile(bootstraps[,i], 0.05))

  }

###############################
###############################
# Now the plot

newdata$Predominant_habitat <- factor(newdata$Predominant_habitat, levels = c("Primary vegetation", "Secondary vegetation", "Plantation forest", "Pasture", "Cropland", "Urban"))

levels(newdata$Predominant_habitat) <- c("PriVeg","SecVeg","PlnFor","Pas","Crp","Urb")


for(r in unique(realms)){

  p_d <- newdata %>% dplyr::filter(Realm == r)
  
  pri_intercept <- p_d %>% dplyr::filter(Predominant_habitat == "PriVeg") %>% pull(predicted)

  
  pd <- position_dodge(0.5)
plot <- ggplot(data = p_d, aes(x = Predominant_habitat, y = predicted, colour = Predominant_habitat)) +
  geom_errorbar(aes(ymin=ifelse(lower< 0, 0, lower) , ymax= upper), colour="black", width=.1, position=pd, linetype = 1) +
  geom_point(position=pd,size=6)+
  xlab("Land use") +
  ylab("Holeyness") +
  scale_colour_hue(name="Use intensity",    # Legend label, use darker colors
                   breaks=c("Minimal", "Light","Intense","All"),
                   labels=c("Minimal", "Light","Intense","All"),
                   l=40) +                    # Use darker colors, lightness=40+
  expand_limits(y=0) +
  ggtitle(paste(r))+
  theme_classic() +
  geom_hline(yintercept= pri_intercept, linetype='dotted', col = 'red')+
  theme(legend.justification=c(1,0),
        legend.position=c(1,0.65),
        text = element_text(size = 20))+
  theme(axis.text.x = element_text(angle = 90))

plot(plot)
}
```


These plots at the site level do not show any clear pattern for an increase in the holeyness of functional trait space of communities in land uses of increasing human impact from primary forest to urban across biogeographic realms. This is a shame and can't really be explained especially the fact that urban sites have smaller holes compared to sites in primary forest and other land uses of less human impact. A possible explanation could be that there is functional homogenization in urban sites meaning that functional divergence and richness is reduced leading to the pattern of small holes in urban sites. 

I'm going to possibly look into the roundness of functional trait space to indicate whether there is a significant change in the shape trait space that can reveal just how land use influences the shape and structure of trait space.
